Gravitational Lens Modelling of Flux Ratio Anomalous System B1555+375

### 2015-04-21 papers/B1555_EdgeOn/B1555_letter.tex
Change in Discussion and also new ref

To model B1555+375, we use the lens modeling code {\tt glafic}
\citep{Oguri}.  The inputs to the model are the observed image
positions and flux densities measured by the radio observations of
\citet{Marlow}.  These authors present a singular isothermal ellipsoid
(SIE) model that provides the first fit to the image positions but
does not reproduce the strong flux ratio anomaly between components A
and B.  \citep[see Fig.~6 and Tables 2 \& 3 in][]{Marlow}.  Although
the flux ratio anomaly that is observed in the radio data may be
caused by substructure in the lensing galaxy, our high-resolution near
infrared imaging suggests another explanation.  In both the $K^\prime$
band adaptive optics data and the HST imaging (two rightmost panels in
Fig. 1), the lensing galaxy has a clear edge-on disk component with a
position angle of $\approx 10$ degree.  We therefore model the system
with a disk component in order to see if such a model can explain the
flux ratio anomaly without the need for substructure.  The edge on
disk may also explain the system morphology that is observed in the
near infrared imaging.  In particular, neither image B nor image D is
detected in the Keck AO data (Fig.~\ref{fig:merlin}), and both of
these lensed images lie on or close to the edge-on disk.  Thus, strong
extinction in the disk could cause the lack of detection of these
images.

We have created two separate models of the lensing potential.  In the
first we treat the lens as a SIE bulge and an exponential edge-on disk
(SIE+expdisk).  In the second model the disk is represented as a very
elongated SIE component (2-SIE).  The lens system is red, with a
strong jump in flux between the F814W (roughly I) and F160W (roughly
H) bands (Fig.~\ref{fig:multiband}), so we assign a lens redshift of
$z_\ell = 1.0$ and a source redshift of $z_s =1.5$.  The full best-fit parameters are shown in Table
2. Both of these models predict fairly matched results with the radio
observation, and the 2-SIE model has a better performance in flux
ratios. The comparison of observation and our lens modelling is given
in Figure 3 and Tables 3 \& 4.

@@

Gravitational lensing is [one of] the most powerful methods for
studying substructure in galaxies outside of the Local Universe, and
the flux ratio anomaly systems are a key aspect of these
investigations.  However, before using the flux ratio anomaly in any
given system to investigate structure formation and the properties of
dark matter, it is critical to eliminate other possible explanations
for the difference between the observed and expected fluxes.  It
is for this reason that much of the effort in the flux-ratio technique
has been focused on radio-loud lens systems.  At these wavelengths,
dust extinction and microlensing effects should be minimal (although
see Koopmans et al. ?? for a rare example of radio microlensing).
Radio propagation effects can  affect radio fluxes, as is clearly
shown in the B0128 system [REF??].  However, these propagation effects
can be disentangled from substructure lensing because they have well
known dependencies on wavelength. [COMMENTS ABOUT B1555 AND OTHER
RADIO-LOUD SYSTEMS].

@@

The abundance of substructures within galaxy-size haloes has been an undetermined problem in structure formation studies. Discrepencies between numerical simulations and observations within the local universe gives two possible interpretations: an alternative structure formation model or a small sample set from nowaday observations. Difficulties emerge in direct detection when searching beyond the Local Group. Until now, most of known satellites are at high mass end in the mass function. However, the gravitational perturbation from substructures can cause detectable change in surface brightness among lensed images. This makes gravitational lensing a powerful tool for searching substructures. Flux ratio anomalies in strongly lensed quasars become a sign of existence of subhaloes. Total mass of substructures can be derived from their gravitational effects and also give direct measurement of mass function, avoid making assumptions on baryonic matter distribution assoociated with dark matter.

Given the assumption that lens galaxy has smooth gravitational potential, only substructure perturbations can cause flux ratio anomalies (in ideal case, i.e. no dust extinction, micro lensing, scattering etc.). Numerical simulations demostrate this idea with known flux anomalous lensed quasars. In \citet{Xu14}, they found that substructures cannot expalin all radio flux anomalies. By adding subhaloes on to those systems, only one out of six lens systems reproduces the flux anomalies. The rest of them are unlikely to have substructure originated flux anomalies, including our target B1555+375. \citet{Xu14} conclude the failure on B1555+375 is due to improper lens models. This result supports our discovery in both observation and lens modelling, that an edge-on disk lies  between the radio merging double explain its flux ratio anomaly.

From the successful modelling of B1555+375,  we expect more flux anomalous systems requiring complex lens models (e.g. include an exponential disk), rather than just assuming smooth potential. Although the number of radio loud flux anomalous lensed quasars remains small, a new technique using narrow-line emissions will enlarge the sample and bring statistical perspective on substructures in the near future \citep{N14}. If the non-substructure effect from edge-on disks does not included, expected ``false'' substructure detections can serverly bias the data analysis. Even with real substructures presence, the lens galaxy structure can still dominate flux anomalous phenomena. For precise measurements on substurcture population, mass function, and any derived properties from strongly lensed flux anomalous systems, it is crucial to have careful estimation on non-substructure effects. We would like to draw attention to this edge-on disk orginated flux anomalies due to its impacts on substructure searching.

Our results also demostrate that multi-wavelength imaging is helpful to explore flux anomalous lensed quasars. Although radio band provides reliable data on positions and fluxes of lensed images, optical and near-infrared imaging provides more information on mass distribution of the lens galaxy. In the previous paper of SHARP, \citet{SHARP12} show that high resolution IR imaging can help to improve lens modelling. The improvement is even more dramatic in B1555+375. With an high ellipcity edge-on disk, our new lens models adequately explain the flux ratio anomaly. Lack of knowledge of lens galaxies can lead to improper lens modelling and bias to statistics among substructure detections. With the first system analyzing and finding that an edge-on disk from the lens galaxy can dominate its flux ratio anomaly, our follow up will be explore other SHARP systems. In Figure 3, all of these systems have a recognizable edge-on disk in infrared imaging. An edge-on disk structure may also be able to expalin the flux ratio anomalies of these systems.\\

>>>>

To model B1555+375, we use the lens modeling code {\tt glafic}
\citep{Oguri}.  The inputs to the model are the observed image
positions and flux densities measured by the radio observations of
\citet{Marlow}.  These authors present a singular isothermal ellipsoid
(SIE) model that provides the first fit to the image positions but
does not reproduce the strong flux ratio anomaly between components A
and B.  \citep[see Fig.~6 and Tables 2 \& 3 in][]{Marlow}.  Although
the flux ratio anomaly that is observed in the radio data may be
caused by substructure in the lensing galaxy, our high-resolution near
infrared imaging suggests another explanation.  In both the $K^\prime$
band adaptive optics data and the HST imaging (two rightmost panels in
Fig. 1), the lensing galaxy has a clear edge-on disk component with a
position angle of $\approx 10$ degree.  We therefore model the system
with a disk component in order to see if such a model can explain the
flux ratio anomaly without the need for substructure.  The edge-on
disk may also explain the system morphology that is observed in the
near infrared imaging.  In particular, neither image B nor image D is
detected in the Keck AO data (Fig.~\ref{fig:merlin}), and both of
these lensed images lie on or close to the edge-on disk.  Thus, strong
extinction in the disk could cause the lack of detection of these
images.

As a first trail of modelling, we create a single SIE model to 
justify the perforamnce of a simple lensing potential. Like previous studies
\citep{Marlow, Xu14}, flux anomaly in B1555+375 cannot be predcited
with a single elliptical lensing potential. 
We then create two separate models of the lensing potential.  In the
first we treat the lens as a SIE bulge and an exponential edge-on disk
(SIE+expdisk).  In the second model the disk is represented as a very
elongated SIE component (2-SIE).  The lens system is red, with a
strong jump in flux between the F814W (roughly I) and F160W (roughly
H) bands (Fig.~\ref{fig:multiband}), so we assign a lens redshift of
$z_\ell = 1.0$ and a source redshift of $z_s =1.5$.  The full best-fit parameters are shown in Table
2. Both of these models predict fairly matched results with the radio
observation, and the 2-SIE model has a better performance in flux
ratios. The comparison of observation and our lens modelling is given
in Figure 3 and Tables 3 \& 4.

@@

Gravitational lensing is [one of] the most powerful methods for
studying substructure in galaxies outside of the Local Universe, and
the flux ratio anomaly systems are a key aspect of these
investigations.  However, before using the flux ratio anomaly in any
given system to investigate structure formation and the properties of
dark matter, it is critical to eliminate other possible explanations
for the difference between the observed and expected fluxes.  It
is for this reason that much of the effort in the flux-ratio technique
has been focused on radio-loud lens systems.  At these wavelengths,
dust extinction and microlensing effects should be minimal (although
see \citet{K2000} for a rare example of radio microlensing).
Radio propagation effects can  affect radio fluxes, as is clearly
shown in the B0128 system [REF??, In Dandan's paper?].  However, these propagation effects
can be disentangled from substructure lensing because they have well
known dependencies on wavelength. [COMMENTS ABOUT B1555 AND OTHER
RADIO-LOUD SYSTEMS, from K03 paper].

@@

Recent studies on numerical simulation support our discover on B1555+375. In \citet{Xu14}, they found that substructures cannot expalin all radio flux anomalous phenomena in radio loud lensed systems. By adding subhaloes on to those systems, only one out of six lens systems reproduces the flux anomalies. The rest of them are unlikely (only few percent likelihood) to have substructure originated flux anomalies, including our target B1555+375. \citet{Xu14} conclude the failure in predicting fluxes of those lens systems is due to improper lens models. Their results are consistent with the SHARP $K^{\prime}$ band observations, which in B1555+375, an edge-on disk lies cross the radio merging double. Here we would like to draw attention to the possible impacts on substructure searching from the enge-on disk orginated flux anomalies. In SHARP observations, we find B1555+375 is not the only system having edge-on disk detection in high resolution near-infrared imaging (see Fig 4). It is a sign that researchers should be careful about the flux anomalies fraction in lensed quasars. It may be worth considering an upper limit of flux anomalies fraciton from observation results to avoid over estimate the number of substructure detections. Although the number of radio loud flux anomalous lensed quasars remains small, a new technique using narrow-line emissions will enlarge the sample and bring statistical perspective on substructures in the near future \citep{N14}. For further analysis on substurcture population, mass function etc., it is always important to rule out other non-substructure effects.

In this paper we also demostrate the power of combining multiwavelength high resolution data. All the flux anomalous lensed quasars are radio loud systems and are first discovered in radio observations. Radio data provides reliable information on positions and fluxes of lensed images but often unable to detect lens galaxies. Since lens galaxies are mostly in low redshift, optical and near-infrared observations are ideal to measure the light distribution of lens gaxies and obtain prior on underlying mass distribution. In the previous paper of SHARP, \citet{SHARP12} show that high resolution IR imaging can help to improve lens modelling. The improvement is even more dramatic in B1555+375. With HST and SHARP AO high resolution imaging, we detect an edge-on disk which does not appear in any previous discussions on B1555+375. This new discovered edge-on disk provides a strong hint to our lens modelling and we show that it adequately explain sthe flux ratio anomaly in B1555+375. Our results demostrate that lack of knowledge of lens galaxies can lead to improper lens models. Modelling with only four image positions from radio data can also cause a lot of degeneracy. With the first system analyzing and finding an edge-on disk originated flux ratio anomaly, our follow up will be explore other similar SHARP systems. In Figure 3, all of these systems have a recognizable edge-on disk in infrared imaging. An edge-on disk structure may also be able to expalin the flux ratio anomalies of these systems.\\

%From the successful modelling of B1555+375,  we expect more flux anomalous systems requiring complex lens models (e.g. include an exponential disk), rather than just assuming smooth potential. Although the number of radio loud flux anomalous lensed quasars remains small, a new technique using narrow-line emissions will enlarge the sample and bring statistical perspective on substructures in the near future \citep{N14}. If the non-substructure effect from edge-on disks does not included, expected ``false'' substructure detections can serverly bias the data analysis. Even with real substructures presence, the lens galaxy structure can still dominate flux anomalous phenomena. For precise measurements on substurcture population, mass function, and any derived properties from strongly lensed flux anomalous systems, it is crucial to have careful estimation on non-substructure effects. We would like to draw attention to this edge-on disk orginated flux anomalies due to its impacts on substructure searching.
